**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-01-05 Schloss_PacBio16S_PeerJ_2016.Rmd
Update code and text for Identifying correlates of increased sequencing error section

mock_error_basic <- mock_error_good[mock_error_good$good_coords & mock_error_good$good_homop,]

@@

```{r initialError}

```

```{r barcodePrimerAnalysis}
getFragment_BPErrorRates <- function(folder){
	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder, ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]
	bc.primer <- composite$fbdiffs + composite$rbdiffs + composite$fpdiffs + composite$rpdiffs

	return(cbind(composite$error, bc.primer))
}

errors <- lapply(regions, getFragment_BPErrorRates)

errors.table <- matrix(errors[[1]], ncol=2, byrow=F)
err.bcprimer <- aggregate(100*errors.table[,1], by=list(errors.table[,2]),
													function(x){c(mean(x), length(x))})

getBCPrimerError <- function(folder){
	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder,
																		 ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]
	mismatch <- composite$fbdiffs + composite$rbdiffs + composite$fpdiffs + composite$rpdiffs

	good0 <- mismatch == 0
	composite.good0 <- composite[good0,]

	good1 <- mismatch <= 1
	composite.good1 <- composite[good1,]

	return(c(mean(composite.good0[,"error"]), nrow(composite.good0)/nrow(composite),
	mean(composite.good1[,"error"]), nrow(composite.good1)/nrow(composite) ))
}

bcprimer.error <- matrix(unlist(lapply(regions, getBCPrimerError)), nrow=6, byrow=T)
rownames(bcprimer.error) <- regions
colnames(bcprimer.error) <- c("0.error", "0.frac", "1.error", "1.frac")

#         0.error    0.frac     1.error    1.frac
# v4  0.002616919 0.6417469 0.003711755 0.9081720
# v13 0.011102712 0.5799452 0.012493713 0.8675194
# v35 0.010838504 0.6464069 0.011901910 0.8934573
# v15 0.005375053 0.5079055 0.007029568 0.8338150
# v16 0.010374438 0.5456238 0.011893323 0.8609559
# v19 0.008543379 0.5072207 0.010086351 0.8237548
```
```{r coverageAnalysis}
errorAtCoverageCutoff <- function(folder){
	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder, ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]

	total.seqs <- nrow(composite)

	deep <- composite$freq >= 10
	composite <- composite[deep,]

	covered.seqs <- nrow(composite)

	error <- sum(composite$mismatches)/sum(composite$total)
	frac.kept <- covered.seqs / total.seqs
	return(c(error=error, kept=frac.kept))
}

coverage.error <- unlist(lapply(regions, errorAtCoverageCutoff))
coverage.error <- matrix(coverage.error, ncol=2, byrow=T)
rownames(coverage.error) <- regions
colnames(coverage.error) <- c("10.error", "10.frac")

#       10.error   10.frac
#v4  0.002524566 0.7897874
#v13 0.010359113 0.6592051
#v35 0.010084961 0.6642832
#v15 0.006534609 0.6717103
#v16 0.010039601 0.4605835
#v19 0.009506114 0.5620395

coverage.reduction <- (1-coverage.error[,"10.error"]/basic.error)*100
#      v4      v13      v35      v15      v16      v19
#53.02015 29.73870 26.54859 27.61302 29.54498 27.03110
```
```{r qualityScoreAnalysis}
getErrorRateFromAveQ <- function(folder, threshold=60){
	write(folder, "")
	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder, ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]

	rates <- composite[composite$aveQ >= threshold,]

	remaining.error <- mean(rates[,"error"])
	fraction.kept <- nrow(rates)/nrow(composite)

	return(c(remaining.error, fraction.kept))
}

aveq.error <- matrix(unlist(lapply(regions, getErrorRateFromAveQ)), ncol=2, byrow=T)
rownames(aveq.error) <- regions
colnames(aveq.error) <- c("60.error", "60.frac")

#       60.error   60.frac
#v4  0.002623098 0.8692667
#v13 0.009914371 0.7608497
#v35 0.009035060 0.7697533
#v15 0.004553350 0.7005270
#v16 0.008921902 0.6561142
#v19 0.005714256 0.5010315

quality.reduction <- (1-aveq.error[,"60.error"]/basic.error)*100
```
```{r oligosCoverageAnalysis}
oligosCoverage <- function(folder){
	write(folder, "")
	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder, ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]

	mismatch <- composite$fbdiffs + composite$rbdiffs + composite$fpdiffs + composite$rpdiffs
	good0 <- mismatch == 0 & composite$freq >= 10
	composite.good0 <- composite[good0,]

	good1 <- mismatch <= 1 & composite$freq >= 10
	composite.good1 <- composite[good1,]

	return(c(mean(composite.good0[,"error"]), nrow(composite.good0)/nrow(composite),
	mean(composite.good1[,"error"]), nrow(composite.good1)/nrow(composite) ))
}

oligosCoverage.error <- matrix(unlist(lapply(regions, oligosCoverage)), ncol=4, byrow=T)
rownames(oligosCoverage.error) <- regions
colnames(oligosCoverage.error) <- c("0.10.error", "0.10.frac", "1.10.error", "1.10.frac")

#     0.10.error 0.10.frac  1.10.error 1.10.frac
#v4  0.001565052 0.5567846 0.002068184 0.7513111
#v13 0.009371691 0.4353586 0.009899256 0.6180905
#v35 0.009096211 0.4907043 0.009503092 0.6321058
#v15 0.004050098 0.3803978 0.005253236 0.5899354
#v16 0.007945951 0.2960894 0.008956478 0.4301676
#v19 0.006706336 0.3188918 0.007903900 0.4927793
```
```{r oligosQScoreAnalysis}
oligosQScore <- function(folder){
	write(folder, "")
	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder, ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]

	mismatch <- composite$fbdiffs + composite$rbdiffs + composite$fpdiffs + composite$rpdiffs

	good <- mismatch == 0 & composite$aveQ >= 60
	composite.good0 <- composite[good,]

	good1 <- mismatch <= 1 & composite$aveQ >= 60
	composite.good1 <- composite[good1,]

	return(c(mean(composite.good0[,"error"]), nrow(composite.good0)/nrow(composite),
	mean(composite.good1[,"error"]), nrow(composite.good1)/nrow(composite) ))
}

oligosQScore.error <- matrix(unlist(lapply(regions, oligosQScore)), ncol=4, byrow=T)
rownames(oligosQScore.error) <- regions
colnames(oligosQScore.error) <- c("0.60.error", "0.60.frac", "1.60.error", "1.60.frac")

#     0.60.error 0.60.frac  1.60.error 1.60.frac
#v4  0.001715135 0.6015066 0.002215499 0.8232097
#v13 0.009153895 0.5013705 0.009700915 0.7108269
#v35 0.008365403 0.5557740 0.008690839 0.7281015
#v15 0.003217812 0.4102346 0.003993206 0.6282727
#v16 0.007589961 0.4115456 0.008358922 0.6101800
#v19 0.004870705 0.3209549 0.005294429 0.4653699
```
```{r coverageQScoreAnalysis}
coverageQScore <- function(folder){
	write(folder, "")
		composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder, ".composite", sep=""), header=T, row.names=1)
		composite <- composite[composite$reason == "g",]

	good <- composite$freq >= 10 & composite$aveQ>= 60
	composite.good <- composite[good,]

	return(c(mean(composite.good[,"error"]), nrow(composite.good)/nrow(composite) ))
}

coverageQScore.error <- matrix(unlist(lapply(regions, coverageQScore)), ncol=2, byrow=T)
rownames(coverageQScore.error) <- regions
colnames(coverageQScore.error) <- c("10.60.error", "10.60.frac")

#    10.60.error 10.60.frac
#v4  0.002361882  0.7874988
#v13 0.010016818  0.6541800
#v35 0.009296289  0.6564176
#v15 0.004481083  0.5824549
#v16 0.009612120  0.4587213
#v19 0.005692918  0.3884468
```
```{r allFiltersAnalysis}
allFilters <- function(folder){
write(folder, "")
		composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder, ".composite", sep=""), header=T, row.names=1)
		composite <- composite[composite$reason == "g",]
	mismatch <- composite$fbdiffs + composite$rbdiffs + composite$fpdiffs + composite$rpdiffs

good0 <- mismatch == 0 & composite$aveQ>= 60 & composite$freq >= 10
composite.good0 <- composite[good0,]

good1 <- mismatch <= 1 & composite$aveQ >= 60 & composite$freq >= 10
composite.good1 <- composite[good1,]

return(c(mean(composite.good0[,"error"]), nrow(composite.good0)/nrow(composite),
mean(composite.good1[,"error"]), nrow(composite.good1)/nrow(composite) ))
}

allFilters.error <- matrix(unlist(lapply(regions, allFilters)), ncol=4, byrow=T)
rownames(allFilters.error) <- regions
colnames(allFilters.error) <- c("0.10.60.error", "0.10.60.frac", "1.10.60.error", "1.10.60.frac")

#    0.10.60.error 0.10.60.frac 1.10.60.error 1.10.60.frac
#v4    0.001515500    0.5562125   0.001993325    0.7499762
#v13   0.009299146    0.4337597   0.009827404    0.6155779
#v35   0.008613200    0.4862353   0.008979207    0.6253128
#v15   0.003078476    0.3488609   0.003905562    0.5268616
#v16   0.007738039    0.2954687   0.008814923    0.4295469
#v19   0.004797594    0.2496316   0.005266467    0.3631005
```

***Identifying correlates of increased sequencing error.***
In contrast to the 454 and Illumina-based platforms where the sequencing
quality decays with length, the consensus sequencing approach employed by the
PacBio sequencer is thought to generate a uniform distribution of errors. This
makes it impossible to simply trim sequences to high quality regions. Therefore,
we sought to identify characteristics within sequences that would allow us to
identify and remove those sequences with errors using three different
approaches. First, we hypothesized that errors in the barcode and primer would
be correlated with the error rate for the entire sequence. We observed a strong
relationship between the number of mismatches to the barcodes and primers and the
error rate of the rest of the sequence fragment (Figure 1B). Although
allowing no mismatches to the barcodes and primers yielded the lowest error rate,
that stringent criterion removed a large fraction of the reads from the dataset and
allowing at most one mismatch marginally increased the error rate while preserving
more sequences in the dataset (Figure 2). Second, we hypothesized
that increased sequencing coverage should yield lower error rates. We found that
once we had obtained 10-fold coverage of the fragments, the error rate did not
change appreciably (Figure 1C). When we compared the error
rates of reads with at least 10-fold coverage to those with less coverage, we
reduced the error rate by
`r paste(round(sort(coverage.reduction)[c(1,5)], digits=1), collapse=" to ")`%
for each region except the V4 region for which the error rate was reduced by
`r round(coverage.reduction["v4"], digits=1)`%. Third, based on the earlier
analysis associating errors with quality scores, we used two quality score-based
approaches for identifying reads with errors (Figure 3). We calculated the
minimum average quality score across all 50-nt windows within each sequence and
we also calculated the average quality score across each sequence. We then
associated both methods of calculating the average quality score with the error
rate of the reads and the fraction of sequences that would be retained if each
threshold were selected. Using the sliding window approach we did not observe
any clear break points indicating that one quality score would be better than
another (Figure 3AB). In contrast, using the whole sequence quality score
average we observed a decrease in the error rate and the fraction of sequences
retained when the threshold was increased above 60 (Figure 3CD). When we used
this threshold, we were able to reduce the error rate by
`r paste(round(range(quality.reduction), digits=1), collapse=" to ")`%
(Figure 2A). We noted that the fraction of reads retained
decreased as the length of the fragment increased with retention of
`r round(100*aveq.error["v4", "60.frac"],digits=1)`% of the V4 reads and
`r round(100*aveq.error["v19", "60.frac"],digits=1)`% of the V1-V9 reads (Figure
2B). Next, we asked whether which combinations of culling reads with mismatches
to the expected barcodes and primers, less than 10-fold sequencing coverage, and
an average quality score less than 60 made the most meaningful reductions in the
error rate while preserving the most reads when implemented with the basic
curation pipeline (Figure 2B). We observed similar error rates when we required
one or fewer mismatches to the barcodes and primers and an average quality score
above 60 as when we also required a minimum 10-fold coverage. Culling sequences
that had more than one mismatch to the barcodes and primers and those with an
average quality score less than 60 reduced the error rate to between
`r paste(round(100*range(oligosQScore.error[,"1.60.error"]), digits=2), collapse=" and ")`.
This procedure resulted in the removal of
`r paste(rev(round(100*(1-range(oligosQScore.error[,"1.60.frac"])), digits=0)), collapse=" and ")`%
of the reads (Figure 2). The remainder of this paper uses this sequence curation
approach.

>>>>

mock_error_basic <- mock_error_good[mock_error_good$good_coords & mock_error_good$good_homop,]
basic_n <- aggregate(mock_error_basic$error, by=list(mock_error_basic$region), length)$x
names(basic_n) <- names(basic_rates)

@@

```{r barcode_primer_analysis}
mock_error_basic$pbdiffs <- mock_error_basic$fbdiffs + mock_error_basic$rbdiffs + mock_error_basic$fpdiffs + mock_error_basic$rpdiffs

#This is for the Figure 2B plot...
pb_overall <- aggregate(mock_error_basic$error, by=list(mock_error_basic$pbdiffs), function(x){c(mean=mean(100*x))})

# want to get the error rate and fraction of sequences for each region with a
# threshold of mismatches or fewer

pb_by_region_mm <- aggregate(mock_error_basic$error, by=list(mock_error_basic$pbdiffs, mock_error_basic$region), function(x){c(mean=mean(100*x), nseqs=length(x))})

aggregate_pb_data <- function(region){
	pb_region <- pb_by_region_mm[pb_by_region_mm$Group.2 == region,]

	product <- pb_region$x[,"mean"] * pb_region$x[,"nseqs"]
	pb_region_nseqs <- cumsum(pb_region$x[,"nseqs"])
	pb_region_error <- cumsum(product) / pb_region_nseqs
	pb_region_frac <- pb_region_nseqs / sum(pb_region$x[,"nseqs"])
	c(error=pb_region_error, fraction=pb_region_frac)
}

pb_data <- sapply(regions, aggregate_pb_data)

pb_error <- pb_data[grep("error", rownames(pb_data)),]
rownames(pb_error) <- 0:6

pb_fraction <- pb_data[grep("fraction", rownames(pb_data)),]
rownames(pb_fraction) <- 0:6

pb_error_reduction_0 <- 100 * (1 - pb_error["0",]/basic_rates)
pb_error_reduction_1 <- 100 * (1 - pb_error["1",]/basic_rates)

pb_nseqs_reduction_0 <- 100 * (1 - pb_fraction["0",])
pb_nseqs_reduction_1 <- 100 * (1 - pb_fraction["1",])
```

```{r coverage_analysis}
#This is for the Figure 2C plot...
coverage_overall_agg <- aggregate(100 * mock_error_basic$error, by=list(mock_error_basic$coverage), mean)
coverage_overall <- coverage_overall_agg$x
names(coverage_overall) <- coverage_overall_agg$Group.1
coverage_overall <- coverage_overall[as.character(3:40)]

# need to get the error rate for coverage values between 3 and 40 for each region

error_by_coverage <- function(region){
	coverage_region <- mock_error_basic[mock_error_basic$region == region,]
	coverage_error_agg <- aggregate(100 * coverage_region$error, by=list(coverage_region$coverage), mean)
	coverage_error <- coverage_error_agg[coverage_error_agg$Group.1 <= 40,"x"]
	names(coverage_error) <- coverage_error_agg[coverage_error_agg$Group.1 <= 40,"Group.1"]
	coverage_error
}

coverage_error <- sapply(regions, error_by_coverage)
#seems to plateau right around 10

error_at_below_coverage_cutoff <- function(region, cutoff=10){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	deep <- region_data$coverage >= 10
	coverage_data <- region_data[deep,]

	error <- 100 * sum(coverage_data$mismatches)/sum(coverage_data$total)
	frac_kept <- nrow(coverage_data) / nrow(region_data)
	c(error=error, fraction=frac_kept)
}

coverage_data <- sapply(regions, error_at_below_coverage_cutoff)

coverage_error <- coverage_data["error",]
coverage_fraction <- coverage_data["fraction",]

coverage_error_reduction <- 100 * (1 - coverage_error/basic_rates)
coverage_nseqs_reduction <- 100 * (1 - coverage_fraction)
```

```{r predicted_error_analysis }
#This is for the Figure 2D plot...
pred_overall <- aggregate(100 * mock_error_basic$error, by=list(round(mock_error_basic$pred_error, 3)), mean)

error_at_above_pred_error_cutoff <- function(region, threshold=0.999){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	above_threshold <- region_data[region_data$pred_error >= threshold,]

	c(error=100*sum(above_threshold$mismatches)/sum(above_threshold$total), nseqs=nrow(above_threshold)/nrow(region_data))
}

pred_error_data <- sapply(regions, error_at_above_pred_error_cutoff, threshold=0.999)

pred_error_reduction <- 100 * (1 - pred_error_data["error",]/basic_rates)
pred_nseqs_reduction <- 100 * (1 - pred_error_data["nseqs",])
```

```{r oligos_coverage_analysis}
get_oligos_coverage <- function(region){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	good0 <- region_data$pbdiffs <= 0 & region_data$coverage >= 10
	region_data_good0 <- region_data[good0,]

	good1 <- region_data$pbdiffs <= 1 & region_data$coverage >= 10
	region_data_good1 <- region_data[good1,]

	c(	100*mean(region_data_good0[,"error"]), nrow(region_data_good0)/nrow(region_data),
		100*mean(region_data_good1[,"error"]), nrow(region_data_good1)/nrow(region_data))
}

oligos_coverage_error <- matrix(unlist(lapply(regions, get_oligos_coverage)), ncol=4, byrow=T)
rownames(oligos_coverage_error) <- regions
colnames(oligos_coverage_error) <- c("0.10.error", "0.10.frac", "1.10.error", "1.10.frac")

#     0.10.error 0.10.frac 1.10.error 1.10.frac
#V1V3  0.4515344 0.5898447  0.4632910 0.7968639
#V1V5  0.4187311 0.5557715  0.4270938 0.7384810
#V1V6  0.4399351 0.5410259  0.4460726 0.7167325
#V1V9  0.4317629 0.4444776  0.4420337 0.5752849
#V3V5  0.6421004 0.6854427  0.6555240 0.8102832
#V4    0.6825316 0.5512332  0.7023460 0.8245533

oligos_coverage_error_reduction <- 100 * (1 - oligos_coverage_error[,"1.10.error"]/basic_rates)
oligos_coverage_nseqs_reduction <- 100 * (1 - oligos_coverage_error[,"1.10.frac"])
```

```{r oligosQScoreAnalysis}
get_oligos_pred_error <- function(region){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	good0 <- region_data$pbdiffs <= 0 & region_data$pred_error >= 0.999
	region_data_good0 <- region_data[good0,]

	good1 <- region_data$pbdiffs <= 1 & region_data$pred_error >= 0.999
	region_data_good1 <- region_data[good1,]

	c(	100 * mean(region_data_good0[,"error"]), nrow(region_data_good0)/nrow(region_data),
		100 * mean(region_data_good0[,"error"]), nrow(region_data_good0)/nrow(region_data) )
}

oligos_pred_error <- matrix(unlist(lapply(regions, get_oligos_pred_error)), ncol=4, byrow=T)
rownames(oligos_pred_error) <- regions
colnames(oligos_pred_error) <- c("0.999.error", "0.999.frac", "1.999.error", "1.999.frac")

#     0.999.error 0.999.frac 1.999.error 1.999.frac
#V1V3   0.2806049  0.4749822   0.2806049  0.4749822
#V1V5   0.2177965  0.4266279   0.2177965  0.4266279
#V1V6   0.2168503  0.4296550   0.2168503  0.4296550
#V1V9   0.2177073  0.3438720   0.2177073  0.3438720
#V3V5   0.2229757  0.5131411   0.2229757  0.5131411
#V4     0.3485654  0.3961328   0.3485654  0.3961328

oligos_pred_error_reduction <- 100 * (1 - oligos_pred_error[,"1.999.error"]/basic_rates)
oligos_pred_nseqs_reduction <- 100 * (1 - oligos_pred_error[,"1.999.frac"])
```

```{r coverage_pred_error_analysis}
get_coverage_pred <- function(region){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	good <- region_data$coverage >= 10 & region_data$pred_error >= 0.999
	region_data_good <- region_data[good,]

	c(100 * mean(region_data_good[,"error"]), nrow(region_data_good)/nrow(region_data) )
}

coverage_pred_error <- matrix(unlist(lapply(regions, get_coverage_pred)), ncol=2, byrow=T)
rownames(coverage_pred_error) <- regions
colnames(coverage_pred_error) <- c("10.999.error", "10.999.frac")

#    10.999.error 10.999.frac
#V1V3   0.2843963    65.86436
#V1V5   0.2188792    56.55108
#V1V6   0.2178326    57.33756
#V1V9   0.2191284    42.66675
#V3V5   0.2236951    60.74254
#V4     0.3527900    61.79213

coverage_pred_error_reduction <- 100 * (1 - coverage_pred_error[,"10.999.error"]/basic_rates)
coverage_pred_nseqs_reduction <- 100 * (1 - coverage_pred_error[,"10.999.frac"])
```

```{r all_filters_analysis}
get_all_filters <- function(region){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	good0 <- region_data$pbdiffs == 0 & region_data$pred_error >= 0.999 & region_data$coverage >= 10
	region_data_good0 <- region_data[good0,]

	good1 <- region_data$pbdiffs <= 1 & region_data$pred_error >= 0.999 & region_data$coverage >= 10
	region_data_good1 <- region_data[good1,]

	c(	100 * mean(region_data_good0[,"error"]), nrow(region_data_good0)/nrow(region_data),
		100 * mean(region_data_good1[,"error"]), nrow(region_data_good1)/nrow(region_data) )
}

all_filters_error <- matrix(unlist(lapply(regions, get_all_filters)), ncol=4, byrow=T)
rownames(all_filters_error) <- regions
colnames(all_filters_error) <- c("0.10.999.error", "0.10.999.frac", "1.10.999.error", "1.10.999.frac")

#    0.10.60.error 0.10.60.frac 1.10.60.error 1.10.60.frac
#v4    0.001515500    0.5562125   0.001993325    0.7499762
#v13   0.009299146    0.4337597   0.009827404    0.6155779
#v35   0.008613200    0.4862353   0.008979207    0.6253128
#v15   0.003078476    0.3488609   0.003905562    0.5268616
#v16   0.007738039    0.2954687   0.008814923    0.4295469
#v19   0.004797594    0.2496316   0.005266467    0.3631005

#     0.10.999.error 0.10.999.frac 1.10.999.error 1.10.999.frac
#V1V3      0.2808218     0.4696170      0.2840387     0.6183602
#V1V5      0.2184805     0.4192252      0.2182989     0.5374066
#V1V6      0.2170989     0.4182253      0.2174090     0.5421862
#V1V9      0.2176989     0.3213802      0.2191334     0.4016057
#V3V5      0.2241187     0.5038275      0.2240779     0.5841414
#V4        0.3492870     0.3932393      0.3526755     0.5665274

all_filters_error_reduction <- 100 * (1 - all_filters_error[,"1.10.999.error"]/basic_rates)
all_filters_nseqs_reduction <- 100 * (1 - all_filters_error[,"1.10.999.frac"])
```

```{r summary}
base_error_reduction <- rbind(	pb_error_reduction_0,
								pb_error_reduction_1,
								pred_error_reduction,
								coverage_error_reduction,
								coverage_pred_error_reduction,
								oligos_coverage_error_reduction,
								oligos_pred_error_reduction,
								all_filters_error_reduction)

base_nseqs_reduction <- rbind(	pb_nseqs_reduction_0,
								pb_nseqs_reduction_1,
								pred_nseqs_reduction,
								coverage_nseqs_reduction,
								coverage_pred_nseqs_reduction,
								oligos_coverage_nseqs_reduction,
								oligos_pred_nseqs_reduction,
								all_filters_nseqs_reduction)
```

***Identifying correlates of increased sequencing error.***
In contrast to the 454 and Illumina-based platforms where the sequencing
quality decays with length, the consensus sequencing approach employed by the
PacBio sequencer is thought to generate a uniform distribution of errors. This
makes it impossible to simply trim sequences to high quality regions. Therefore,
we sought to identify characteristics within sequences that would allow us to
identify and remove those sequences with errors using three different
approaches. First, we hypothesized that errors in the barcode and primer would
be correlated with the error rate for the entire sequence. We observed a strong
relationship between the number of mismatches to the barcodes and primers and the
error rate of the rest of the sequence fragment (Figure 1B). Although
allowing no mismatches to the barcodes and primers yielded the lowest error rate,
that stringent criterion removed a large fraction of the reads from the dataset.
Allowing at most one mismatch only marginally increased the error rate while
retaining more sequences in the dataset (Figure 2). Second, we hypothesized that
increased sequencing coverage should yield lower error rates. We found that once
we had obtained 10-fold coverage of the fragments, the error rate did not change
appreciably (Figure 1C). When we compared the error rates of reads with at least
10-fold coverage to those with less coverage, we reduced the error rate by
`r paste(format(range(coverage_reduction), digits=3, nsmall=1), collapse=" to ")`%
(Figure 2). Third, based on the earlier analysis finding a relationship between the
predicted error rate and the observed error rate, we sought a minimum predicted
error rate that would allow us to reduce the observed error rate. The average
observed error rate for sequences with predicted error rates between 0.01 and
0.10% was linear (Figure 1D). We decided to use a threshold of 0.01% because a
large number of sequence reads were lost when we used a smaller threshold. When
we used this threshold, we were able to reduce the error rate by
`r paste(format(range(pred_error_reduction), digits=3, nsmall=1), collapse=" to ")`%
(Figure 2A). Finally, we quantified the effect of combining filters. We found that
any combination of filters that included the predicted error rate threshold had
the most significant impact on reducing the observed error rate. Furthermore,
the inclusion of the mismatch and coverage filters had a negligible impact on
error rates, but had a significant impact on the number of sequences included in
the analysis. For instance, among the V1-V9 data, requiring sequences to have a
predicted error rate less than 0.01% resulted in a `r format(base_error_reduction["pred_error_reduction","V1V9"], digits=3, nsmall=1)`% and resulted in the removal of `r format(base_nseqs_reduction["pred_error_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences. Adding the mismatch or coverage filter
had no effect on the reduction of error, but resulted in the removal of
`r format(base_nseqs_reduction["oligos_pred_nseqs_reduction","V1V9"], digits=3, nsmall=1)` and
`r format(base_nseqs_reduction["coverage_pred_nseqs_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences, respectively. Use of all filters had no impact on the reduction in the observed
error rate, but resulted in the removal of `r format(base_nseqs_reduction["all_filters_nseqs_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences. The remainder of this paper only uses sequences
with a predicted error rate less than 0.01%.
